---
title: "OSM_read_timelapse"
author: "Marissa Dyck"
date: "2024-04-03"
output: 
  html_document:
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Before you begin

## Notes

A few notes about this script.

If you are running this with the 2022-2023 data make sure you download the whole (OSM_2022-2023 GitHub repository)[https://github.com/ACMElabUvic/OSM_2022-2023] from the ACMElabUvic GitHub. This will ensure you have all the files, data, and proper folder structure you will need to run this code and associated analyses.

Also make sure you open RStudio through the R project (OSM_2022-2023.Rproj) this will automatically set your working directory to the correct place (wherever you saved the repository) and ensure you don't have to change the file paths for some of the data. 

If you have question please email the most recent author, currently   

Marissa A. Dyck   
Postdoctoral research fellow    
University of Victoria    
School of Environmental Studies     
Email: [marissadyck17@gmail.com](marissadyck17@gmail.com)     

## Netdrive access

This script relys on the user having access to the ACME lab Netdrive (you can view the .html output of this file if you don't have access and just want to see what the script did).

Helpful instructions for connecting to and navigating the Netdrive can also be found here: [https://docs.google.com/document/d/1Z72IrlIXO8MUHCoVztcMrMdL10R2tHrHThEgfow1Cu0/edit ](https://docs.google.com/document/d/1Z72IrlIXO8MUHCoVztcMrMdL10R2tHrHThEgfow1Cu0/edit).


## Install packages

If you don't already have the following packages installed, use the code below to install them. *NOTE this will not run automatically as eval=FALSE is included in the chunk setup (i.e. I don't want it to run everytime I run this code since I have the packages installed)


```{r install packages, eval=FALSE}

install.packages(tidyverse) 
install.packages(withr) 
```


## Load libraries

Then load the packages to your library.

```{r libraries}

library(tidyverse) # data tidying, visualization, and much more; this will load all tidyverse packages, can see complete list using tidyverse_packages()
library(withr) # used to temporarily set wd
```




# Timelapse data

## Import data files

First, we need to temporarily set your working drive to the ACME lab Netdrive to import the Timelapse data files (there are too many to efficiently store on GitHub or GoogleDrive each year). We will use the `with_dir()` function in the *withr* package to do this.

The code below will read in all the files with a .csv extension from any subfolder/s within the main folder provided in the `path =` argument. I have also specified for it to ignore certain subfolders (e.g. backups). Then it will name each file based on the base name of the .csv file.

This only works to import the timelapse csv files **if** they all have unique file names, are found within the subfolders of the main path folder, and there aren't extraneous files with a .csv extension that we don't want to import. 

If the Netdrive folder structure changes in future years this code *may* not work or *may* need to be adapted to ignore additional subfolders


> this takes a while to run because R is looking in every sub-folder at every file to see if it matches the extension (i.e. .csv) and then saving the file name and path and then going back and removing the files from the 'backups' sub-folders. Maybe not the most elegant way to do this but beats having to manually copy and paste each file individually.

```{r import, eval=FALSE}
# temporarily set the working directory to import from the NetDrive
with_dir(new = '/Volumes/acmelab/1.Resources/2.Arrays/Alberta/OSM/2022-2023',
         
# make a list of the .csv file names from the sub-folders within the '1. Imagery/OSM_2022_LU01' folder
OSM_2022_timelapse_files <- list.files(
  path = '1. Imagery',
  recursive = TRUE,
  pattern = "\\.*csv$",
  full.names = TRUE) %>% 

  
  # exclude files from backups sub-folders
  stringr::str_subset(.,
                      'Backups',
                      negate = TRUE) %>% 
  
  # set names and read csv files for remaining files in using purrr (map)
  {setNames(map(., read_csv), 
            sub("\\.csv$", 
                "", 
                basename(.)))})
  
  

  
```
IT WORKS!

## Export data files

Now we need to save these files with their list name as the file name in the new folder. 

Again we will need to temporarily set the working drive to the netdrive location we want to save the csv files in for this step 
```{r export data, eval=FALSE}

# set the working directory again to the netdrive
with_dir(new = '/Volumes/acmelab/1.Resources/2.Arrays/Alberta/OSM/2022-2023',
         
         # use purrr::imap to save all files to new folder with file names from list
         purrr::imap(
           OSM_2022_timelapse_files,
           ~write_csv(.x,
                      file = paste0('2. Timelapse Files/',
                                    .y,
                                    '.csv'))))

```

